Customizing Software by Direct Manipulation of Tabular Data

### 2020-05-22 paper.tex
add more dnj edits
new example section

We have developed a browser extension called Wildcard which uses web
scraping techniques to implement table-driven customization for existing
Web applications. In Section~\ref{sec:examples}, we introduce the key
ideas of table-driven customization by presenting several examples of
real customizations implemented in Wildcard.

@@

\hypertarget{sec:examples}{%
\section{Examples}\label{sec:examples}}

\emph{todo: maybe redo this as a separate example to avoid self
plagiarism?}

To concretely illustrate the end user experience of table-driven
customization, here are several real examples of using the Wildcard
browser extension to customize websites.

\begin{figure*}
\hypertarget{fig:airbnb-demo}{%
\centering
\includegraphics{media/airbnb-demo-300dpi.png}
\caption{Using Wildcard to augment the Airbnb search page for booking
accommodations}\label{fig:airbnb-demo}
}
\end{figure*}

\hypertarget{augmenting-search-results}{%
\subsection{Augmenting search results}\label{augmenting-search-results}}

In 2012, the travel site Airbnb removed the ability to sort
accommodation searches by price. Users could still filter by price
range, but could no longer view the cheapest listings first. Many users
complained that the change seemed hostile to users. ``It's so
frustrating! What is the logic behind not having this function?'' said
one user on the
\href{https://community.withairbnb.com/t5/Hosting/Sorting-listing-by-price/td-p/559404}{Airbnb
support forum}. Alas, the feature remains missing to this day.

Using Wildcard, the user can fix this omission, while leaving the page's
design and the rest of its functionality unchanged.{
Figure~\ref{fig:airbnb-demo} shows an overview of augmenting the Airbnb
site.} First, the user opens up the Wildcard panel, which shows a table
corresponding to the search results in the page. As they click around in
the table, the corresponding row in the page is highlighted to indicate
the mapping between the views.

Then, the user clicks on the price column header to sort the spreadsheet
and the Airbnb UI by price{ (Figure~\ref{fig:airbnb-demo}, Note A)}.
They also filter to listings with a user rating above 4.5 (another
feature missing in the original Airbnb UI).

After manipulating the data, the user closes the table view and
continues using the website. Because the application's UI usually has a
nicer visual design than a spreadsheet, Wildcard does not aim to replace
it---at any time, the user can use either the UI, the spreadsheet, or
both together.

Many websites that show lists of data also offer actions on rows in the
table, like adding an item to a shopping cart. Wildcard has the ability
to make these ``row actions'' available in the data table through the
site adapter. In the Airbnb UI, saving multiple listings to a Favorites
list requires tediously clicking through them one by one. Using Wildcard
row actions, the user can select multiple rows and favorite all of them
with a single click{ (Figure~\ref{fig:airbnb-demo}, Note B)}. Similarly,
the user can open the detailed pages for many listings at once.

Next, the user wants to jot down some notes about each listing. To do
this, they type some notes into an additional column in each row, and
the notes appear inside the listings in the original UI{
(Figure~\ref{fig:airbnb-demo}, Note C)}. The annotations are saved in
the browser and associated with the backend ID of the listing, so they
will appear in future browser sessions that display the same listing.

Wildcard also includes a formula language that enables more
sophisticated customizations. When traveling without a car, it's useful
to evaluate potential places to stay based on how walkable the
surroundings are. Using a formula, the user can integrate Airbnb with
Walkscore, an API that rates the walkability of any location on a 1-100
scale. When the user calls the \texttt{walkscore} formula with the
listing's latitude and longitude as arguments, it returns the walk score
for that location and shows it as the cell value. Because the cell's
contents are injected into the page, the score also appears in the UI{
(Figure~\ref{fig:airbnb-demo}, Note D)}.

\hypertarget{snoozing-todos}{%
\subsection{Snoozing todos}\label{snoozing-todos}}

\begin{figure*}
\hypertarget{fig:airbnb-demo}{%
\centering
\includegraphics{media/todomvc-demo-300dpi.png}
\caption{Using Wildcard to add a "snooze" feature to the TodoMVC todo list app}\label{fig:todomvc-demo}
}
\end{figure*}

In addition to fetching data from other sources, Wildcard formulas can
also perform computations. In this example, the user would like to
augment the TodoMVC todo list app with a ``snooze'' feature, which will
temporarily hide a todo from the list until a certain date.{
Figure~\ref{fig:todomvc-demo} shows an overview of this customization.}

The user opens the table view, which shows the text and completed status
of each todo. They start the customization by adding a new column to
store the snooze date for each todo{ (Figure~\ref{fig:todomvc-demo},
Note A)}.

The next step is to hide snoozed todos. The user creates a
\texttt{snoozed?} column, which uses a formula to compute whether a todo
is snoozed---i.e., whether it has a snooze date in the future{
(Figure~\ref{fig:todomvc-demo}, Note B)}. Then, they simply filter the
table to hide the snoozed todos{ (Figure~\ref{fig:todomvc-demo}, Note
C)}.

Because the built-in \texttt{NOW()} function always returns the current
datetime, snoozed todos will automatically appear once their snooze date
arrives.

Because this implementation of snoozing was built on the spreadsheet
abstraction, it is completely decoupled from this particular todo list
app. We envision that users could share these types of customizations as
generic browser extensions, which could be applied to any site supported
by Wildcard with no additional effort.

\hypertarget{adding-a-custom-datepicker}{%
\subsection{Adding a custom
datepicker}\label{adding-a-custom-datepicker}}

\begin{figure*}
\hypertarget{fig:airbnb-demo}{%
\centering
\includegraphics{media/expedia-demo-300dpi.png}
\caption{Using Wildcard to augment the Expedia page for booking a flight}\label{fig:expedia-demo}
}
\end{figure*}

It might seem that Wildcard is only useful on websites that display
lists of tabular data, but the table metaphor is flexible enough to
represent many types of data. For example, a flight search form on
Expedia can be represented as a single row, with a column corresponding
to each input{ (Figure~\ref{fig:expedia-demo}, Note A)}.

In some of the previous examples, the table cells were read-only
(because users can't, for example, change the name or price of an Airbnb
listing). In this case, the cells are writable, which means that changes
in the table are reflected in the form inputs. This becomes especially
useful when combined with GUI widgets that can edit the value of a table
cell.

Filling in dates for a flight search often requires a cumbersome
workflow: open up a separate calendar app, find the dates for the trip,
and then carefully copy them into the form. In Wildcard, the user can
avoid this by using a datepicker widget that shows the user's personal
calendar events{ (Figure~\ref{fig:expedia-demo}, Note B)}. The user can
directly click on the correct date, and it gets inserted into both the
spreadsheet and the original form.n

@@

To evaluate table-driven customization in practice, we built the
Wildcard browser extension, which implements table-driven customization
in the context of existing websites. It is built in Typescript, and
works across three major browsers: Chrome, Firefox, and Edge.

We have used Wildcard to customize 11 websites, including transactional
sites like Amazon and Uber Eats, and media consumption sites like Hacker
News and Youtube. Table 1 summarizes all these customizations, including
the number of lines of code in the adapter configuration for each site.

@@

\begin{itemize}
\tightlist
\item
  How broad is the range of possible customizations in this paradigm?
\item
  How feasible is it to build DOM scraping adapters for real websites,
  in practice?
\end{itemize}

@@

We have found that table-driven customization can serve a broad range of
useful purposes. Here we expand on some examples that illuminate various
aspects of using the system in practice.

@@

In other cases we have found that a lack of sorting options seems less
intentional, and simply a missing feature. For example, the Instacart
grocery delivery service doesn't allow users to sort their grocery cart
by price or category. Our guess is that they simply haven't gotten
around to implementing the feature yet.

In the current implementation of Wildcard, users can only sort and
filter entries which are shown on the current page, which means that
users are not entirely liberated from the suggestions of the opaque
algorithm. We could work around this restriction in the future by
fetching content across multiple pages, but we've also found that
sorting/filtering a single page of a paginated list is often an
acceptable outcome, and even a preferable one. For example, it's more
useful to sort 30 recommended Youtube videos than to try to sort all
videos on Youtube; similarly it's convenient to view the Hacker News
front page sorted by read time, rather than all Hacker News articles.

@@

Many other web annotation systems focus on annotating text or arbitrary
webpage content, but Wildcard only allows for annotating structured
objects extracted by an adapter, which results in quite a different set
of use cases.

Annotating with Wildcard has proven most useful when taking notes on a
list of possible options: e.g., evaluating possible Airbnb locations to
rent. We have also used it with Instacart's online grocery cart, for
jotting down notes like ``should we get more milk?''

@@

Formulas are the most powerful part of the Wildcard system. So far, we
have only built a small number of functions into the language, but by
adding more functions, the fundamental computing model could support a
very broad range of useful computations, as shown by the success of
spreadsheets.

We have found that formulas are useful for fetching data from Web APIs.
We've used them to augment Airbnb listings with walkability scores, and
augment Hacker News articles with estimated read times. One challenge of
the current language design is that supporting each web API requires
adding a new function to the language using code, because web APIs
typically return complex JSON data structures which don't make sense to
display in a single table cell. Adding functions for handling JSON data
to the formula language might help resolve this.

@@

``Cell editors'' are UI widgets that expose a custom editing UI for a
single cell of the table view. A programmer building a cell editor only
needs to integrate it with the table viewer; actually propagating values
into the website UI is handled by the site-specific DOM adapter. As a
result, creating a new cell editor can be as easy as just creating glue
code between the table editor and an existing UI widget library.

One benefit of cell editors is that a user can use their personal
information in a web UI without uploading it to the website. To explore
this idea, we created a cell editor based on the FullCalendar Javascript
plugin, which can load data from a Google Calendar. This makes it
convenient to enter dates into a website based on the user's personal
calendar information.

Another benefit of cell editors is that a user can choose their
preferred widget for editing some type of information. We built a cell
editor based on the CKEditor rich text editor. As an example of using
this editor, we integrated it with the Blogger website and used it to
edit blog posts, in place of the Blogger website's built-in editor.

@@

There are many useful customizations which are not possible in the
table-driven customization paradigm. Wildcard can only make
customizations that use the available data exposed in the table. UI
modifications are limited in scope; for example, deleting arbitrary
buttons isn't possible. At one point, we wanted to build an automation
to repeatedly load a grocery delivery website to check for open delivery
slots, but it wasn't clear how to achieve this. Some of these
limitations are specific to the current implementation of the Wildcard
extension, but some are more general to the entire idea of table-driven
customization.

We consider this an acceptable outcome---our goal is simply to make many
useful customizations available to end users with a low threshold, not
to span all possible customizations. Section~\ref{sec:dm} discusses this
point further.

One benefit of Wildcard is \emph{predictability}: once we built a site
adapter for a website, it was generally obvious to us what types of
customizations were and weren't possible, and the UI guided us towards
building customizations that matched the system's model. For example,
it's clear that it's not possible to customize a website using data not
available in the table.

@@

Our second evaluation area is less about the conceptual approach of
table-driven customization, and more about our specific implementation
of customizing existing web applications. In order for third-party
customization through Wildcard to succeed, it is important that creating
usable adapters for existing websites takes minimal effort.

@@

Hutchins, Hollan and Norman \citep{hutchins1985} describe a direct
manipulation interface as one that uses a model-world metaphor, rather
than a conversation metaphor. Rather than conversing with the system
about an assumed world, ``the world is explicitly represented'' and the
user can ``{[}act{]} upon the objects of the task domain themselves.''

Most GUIs today employ direct manipulation, but software customization
tools typically use an imperative programming model, which implements
the conversational metaphor rather than direct manipulation. For
example, in Applescript \citep{cook2007}, the scripting language for
customizing Mac OS applications, here is how a user retrieves a list of
of calendar names from the Calendar application:

@@

Some customization environments (Automator for Mac, Shortcuts for iOS,
Zapier for web APIs) forego text syntax and enable the user to connect
programs and construct automations by dragging and dropping commands.
These environments still do not constitute direct manipulation, though:
the objects being manipulated are in the domain of programming, not the
domain of the task at hand.

@@

However, there is a serious drawback to this approach. MacLean et al
\citep{maclean1990} describe an ideal for user-tailorable systems: a
``gentle slope'' from using to customizing, where small incremental
increases in skill lead to corresponding increments of customization
power. Requiring users to learn programming for customization creates an
abrupt ``cliff,'' where a significant learning investment is required
even to implement simple customizations. Another goal of MacLean et al
is that ``it should be as easy to change the environment as it is to use
it,'' at least for some subset of changes. In scripting languages, even
user-friendly ones like Applescript or Chickenfoot, the experience of
customization does not remotely resemble the experience of use, so these
systems can't meet this goal.

@@

One aspect of directness which we have chosen not to maintain in
Wildcard is enabling customization in the context of the original user
interface, as explored by some other tools like Scotty
\citep{eagan2011}. We have found that augmenting the original UI with an
additional structured representation has numerous benefits. It provides
a consistent experience across all applications, and clearly shows the
user what structured data is available to work with.

Ainsworth et al provide a helpful taxonomy of the value of multiple
representations \citep{ainsworth1999}. In their terms, Wildcard plays a
\emph{complementary role} by supporting a different set of tasks from
the original application, while displaying \emph{shared information}. We
also suspect that Wildcard helps \emph{constructs deeper understanding}
by \emph{subtraction}. By stripping away details and only showing the
essential data in an interface, Wildcard helps users think of new ways
of using that data, outside the specific constraints of the original
application.

It is important to note that there are many specific customizations that
can be achieved in systems like Chickenfoot \citep{bolin2005} that
cannot be reproduced in Wildcard. We consider this an acceptable
tradeoff in exchange for achieving a gentler slope of customization, and
we show in Section~\ref{sec:examples} and Section~\ref{sec:evaluation}
that our model can still implement many useful customizations in
practice. Also, there is sometimes a way to reframe an imperative script
in terms of our direct manipulation model; for example, a script that
iterates through rows in a page adding some additional information could
be reproduced using a formula in Wildcard.

\hypertarget{wrapping-applications-for-customization}{%
\subsection{Wrapping applications for
customization}\label{wrapping-applications-for-customization}}

@@

@@

Table-driven customization is most closely related to other tools that
aim to empower end users to customize software without doing any
programming.

@@

Sifter \citep{huynh2006} enables end users to sort and filter lists of
data based on web scraping, much like Wildcard's sorting features. The
main difference between the systems is that table-driven customization
has many other use cases besides sorting and filtering. Also, Sifter
involves end users in a semi-automated data extraction process, rather
than having programmers create wrappers. This provides coverage of more
websites, but at the expense of complicating the end user experience. We
might explore such techniques in Wildcard in the future, but we think
that it's valuable for end users to have a customization experience free
of web scraping details on supported websites. Finally Sifter implements
scraping across multiple pages, a valuable feature for sorting and
filtering which isn't present in Wildcard.

Thresher \citep{hogue2005} enables end users to create wrappers which
map website content to Semantic Web schemas, making customizations
available using the schema information. We share the general idea of
wrappers with Thresher, but we have chosen to map to a single generic
data type, rather than more specific schemas, which increases the range
of websites and data supported by the system.

There are many software customization tools that offer simplified
versions of programming for end users. Chickenfoot \citep{bolin2005} and
Coscripter \citep{leshed2008} offer friendly syntax for writing web
automation scripts; Applescript \citep{cook2007} attempts a similar goal
for desktop customization. There are visual programming environments for
customization which don't involve writing any text: as a few examples,
\href{https://support.apple.com/guide/automator/welcome/mac}{Automator}
for Mac and
\href{https://apps.apple.com/us/app/shortcuts/id915249334}{Shortcuts}
for iOS are modern options for customizing Apple products, and
\href{https://zapier.com/}{Zapier} enables users to connect different
web applications together visually. As mentioned previously, these tools
all require writing imperative programs in some form; we take a
fundamentally different approach in table-driven customization.

@@

Another relevant area of literature is direct manipulation interfaces
for interacting with structured data. We take much inspiration from
these tools in our work, but apply them in a different domain:
customizing existing software applications, rather than interacting with
databases or constructing software from scratch.

The most closely related work is in systems that offer spreadsheet-like
querying of relational data, as proposed by Liu and Jagadish
\citep{liu2009}. SIEUFERD by Bakke and Karger \citep{bakke2016} is an
example; their paper presents a survey of various other similar tools.
Our work is particularly influenced by the authors' observation that
direct manipulation requires that the user manipulate the \emph{results}
of a database query rather than the query itself, and that the user must
see intermediate results at every step of constructing a computation.
SIEUFERD's interface supports a far wider range of possible queries than
Wildcard, but the basic ideas of the user interface are similar.

@@

Our work is also inspired by the many projects which have explored using
spreadsheets as a foundation for building software applications,
including Object Spreadsheets \citep{mccutchen2016}, Quilt
\citep{benson2014}, Gneiss \citep{chang2014}, Marmite \citep{wong2007},
\href{https://www.glideapps.com/}{Glide}. These projects share the idea
that a spreadsheet is a convenient interface for editing the data
underlying a GUI application. We share that idea, but apply it to
software customization, rather than building software from scratch.

Another system in this space is ScrAPIr, by Alrashed et
al.~\citep{alrashed2020}, which enables end users to access backend web
APIs without programming. ScrAPIr shares a high level goal of end user
empowerment, and also shares the idea of wrappers, by creating a shared
library of wrappers around existing web APIs. However, unlike Wildcard,
ScrAPIr targets explicit APIs exposed by developers, and also doesn't
aim to extend the frontend interfaces of web applications.

>>>>

We have developed a browser extension called Wildcard which uses web
scraping techniques to implement table-driven customization for existing
Web applications. In Section~\ref{sec:examples}, we introduce the key
ideas of table-driven customization through an example scenario.

@@

\hypertarget{sec:examples}{%
\section{Example Scenario}\label{sec:examples}}

To concretely illustrate the end user experience of table-driven
customization, here is a real example of using the Wildcard browser
extension to customize a website.

\begin{itemize}
\tightlist
\item
  Hacker News
\item
  open the table
\item
  sort by comments
\item
  take notes on links
\item
  filter out visited sites with a formula
\item
  add read times with a formula
\end{itemize}

\begin{figure*}
\hypertarget{fig:hacker-news}{%
\centering
\includegraphics[width=\textwidth]{media/hacker-news.png}
\caption{Customizing Hacker News by interacting with a table view}\label{fig:hacker-news}
}
\end{figure*}

@@

To evaluate table-driven customization in practice, we built the
Wildcard browser extension, which implements table-driven customization
in the context of existing websites. It is implemented in Typescript,
and works across three major browsers: Chrome, Firefox, and Edge.

We have used Wildcard to customize 11 websites, including transactional
sites like Amazon and Uber Eats, and media consumption sites like Hacker
News and Youtube. Table 1 summarizes these customizations, showing the
number of lines of code in the adapter configuration for each site.

@@

\begin{itemize}
\tightlist
\item
  How broad is the range of possible customizations in this paradigm?
\item
  How feasible is it to build DOM scraping adapters for real websites?
\end{itemize}

@@

We have found that table-driven customization can serve a broad range of
useful purposes. Here we expand on some archetypal examples that
illuminate various aspects of using the system in practice.

@@

In other cases we have found that a lack of sorting options seems less
intentional, and simply represents a missing feature. For example, the
Instacart grocery delivery service doesn't allow users to sort their
grocery cart by price or category. Our guess is that they simply haven't
gotten around to implementing the feature yet.

In the current implementation of Wildcard, users can only sort and
filter entries that are shown on the current page, which means that
users are not entirely liberated from the suggestions of the opaque
algorithm. We could work around this restriction in the future by
fetching content across multiple pages, but we've also found that
sorting/filtering a single page of a paginated list is often an
acceptable outcome (and sometimes even a preferable one). For example,
it's more useful to sort (by length, say) 30 recommended Youtube videos
than to try to sort all videos on Youtube, or to sort (by read time,
say) the articles on the Hacker News front page than to sort all
non-archived articles.

@@

Many other web annotation systems focus on annotating text or arbitrary
webpage content, but Wildcard limits annotations to structured objects
extracted by an adapter, resulting in quite a different set of use
cases.

Annotating with Wildcard has proven most useful when taking notes on a
list of possible options (e.g., evaluating possible Airbnb locations to
rent). We have also used it with Instacart's online grocery cart, for
jotting down notes like ``should we get more milk?''

@@

Formulas are the most powerful part of the Wildcard system. So far, our
language supports only a small number of predefined functions. Adding
more will allow a very broad range of useful computations, as shown by
the success of spreadsheets.

Formulas are especially useful for fetching data from Web APIs. We've
used them to augment Airbnb listings with walkability scores, and to
augment Hacker News articles with estimated read times. One challenge of
the current language design is that supporting a new web API requires
writing conventional code to add a new function to the language, because
web APIs typically return complex JSON data structures that can't be
easily displayed in a single table cell. Adding functions for handling
JSON data to the formula language might make it possible to access new
APIs without requiring an entire new function in the formula system.

@@

``Cell editors'' are UI widgets that expose a custom editing UI for a
single cell of the table view. A programmer building a cell editor need
only integrate it with the table viewer; propagating values into the
website UI is handled by the site-specific DOM adapter. As a result,
creating a new cell editor can be as easy as just writing some glue code
between the table editor and an existing UI widget library.

An important benefit of cell editors is that they allow users to
incorporate their personal information within the web UI but without
uploading their personal data to the web server. To explore this idea,
we created a cell editor based on the FullCalendar Javascript plugin,
which can load data from a Google Calendar. This makes it convenient to
enter dates into a website based on the user's personal calendar
information.

Another benefit of cell editors is that a user can choose their
preferred widget for editing some type of information. We built a cell
editor based on the CKEditor rich text editor, and demonstrated its use
by integrating it with Google's Blogger website and using it to edit
blog posts, in place of Blogger's built-in editor.

@@

There are many useful customizations that are not currently possible in
the table-driven customization paradigm. Wildcard can only make
customizations that use the available data exposed in the table. UI
modifications are limited in scope; deleting arbitrary buttons isn't
possible, for example. At one point, we wanted to build an automation to
repeatedly load a grocery delivery website to check for open delivery
slots, but it wasn't clear how to achieve this. Some of these
limitations are specific to the current implementation of the Wildcard
extension, but some are more fundamental to the entire idea of
table-driven customization.

We consider this an acceptable outcome; our goal is to make many useful
customizations available to end users with a low threshold, and not to
span all possible customizations. Section 5.1 discusses this point
further.

One benefit of Wildcard is predictability: once we built a site adapter
for a website, it was generally obvious to us what types of
customizations were and weren't possible, and the UI guided us towards
building customizations that matched the system's model. If data is not
available in the table, for example, then it's abundantly clear that
customizations relying on that data will not be expressible.

@@

Our second evaluation area relates less to the conceptual approach of
table-driven customization, and more to the specific implementation of
customizing existing web applications. In order for third-party
customization through Wildcard to succeed, it is important that creating
usable adapters for existing websites takes minimal effort.

@@

Hutchins, Hollan and Norman \citep{hutchins1985} define a direct
manipulation interface as one that uses a model-world metaphor rather
than a conversation metaphor. Instead of presenting an ``assumed'' but
not directly visible world that the user converses with, ``the world is
explicitly represented'' and the user can ``{[}act{]} upon the objects
of the task domain themselves.''

Although most GUIs today employ direct manipulation, software
customization tools typically use an imperative programming model, which
implements the conversational metaphor rather than direct manipulation.
Here, for example, is how a user retrieves a list of of calendar names
from the Calendar application in Applescript \citep{cook2007}, the
scripting language for customizing Mac OS applications:

@@

Some customization environments like Mac Automator and Zapier forego
textual syntax and let the user connect programs and construct
automations by dragging and dropping icons representing commands. These
environments still do not constitute direct manipulation, though: the
objects being manipulated are in the domain of programming, not in the
domain of the task at hand. Imperative programming is a reasonable
choice as the model for building customizations. Turing-complete
programming provides a high ceiling for possible customizations, and a
sequence of commands is a natural fit for automations that simulate a
series of steps taken by the user.

@@

There is, however, a serious drawback to this approach. MacLean et al
\citep{maclean1990} describe an ideal for user-tailorable systems: a
``gentle slope'' from using to customizing, where small incremental
increases in skill lead to corresponding increments of customization
power. Requiring users wanting to customize their applications to learn
programming creates an abrupt ``cliff,'' exacting a significant
investment in learning even to implement the simplest customizations.
Another goal of MacLean et al is to make it ``as easy to change the
environment as it is to use it''---at least for some subset of changes.
But in scripting languages, even user-friendly ones like Applescript or
Chickenfoot, the experience of customization does not remotely resemble
the experience of use, so these systems can't meet this goal.

@@

One aspect of directness that we have chosen not to maintain in Wildcard
is enabling customization in the context of the original user interface,
as explored by tools like Scotty \citep{eagan2011}. We have found that
augmenting the original UI with a distinct, additional representation
provides a consistent experience across all applications, and clearly
shows the user what structured data is available to work with.

Ainsworth et al provide a helpful taxonomy of the value of multiple
representations \citep{ainsworth1999}. In their terms, Wildcard plays a
complementary role by supporting a different set of tasks from the
original application, while displaying shared information. Wildcard may
also help construct deeper understanding by subtraction. By stripping
away details and only showing the essential data in an interface,
Wildcard helps users think of new ways of using that data, outside the
specific constraints of the original application.

It is important to note that there are many specific customizations that
can be achieved in systems like Chickenfoot \citep{bolin2005} that
cannot be reproduced in Wildcard. We consider this an acceptable
tradeoff in exchange for a gentler slope in customization, and we show
in Section~\ref{sec:examples} and Section~\ref{sec:evaluation} that our
model can still implement many useful customizations in practice. Also,
there is sometimes a way to reframe an imperative script in terms of our
direct manipulation model; for example, a script that iterates through
rows in a page adding some additional information can be reproduced
using a formula in Wildcard.

\hypertarget{semantic-wrappers}{%
\subsection{Semantic wrappers}\label{semantic-wrappers}}

@@

\emph{todo: copy over more content from gdoc}

@@

Table-driven customization is most closely related to other tools that
aim to empower end users to customize software without traditional
coding.

@@

Sifter \citep{huynh2006} enables end users to sort and filter lists of
data obtained by web scraping, much like Wildcard's sorting features.
The main difference between the systems is that table-driven
customization has many other use cases besides sorting and filtering.
Also, Sifter involves end users in a semi-automated data extraction
process, rather than having programmers create wrappers. This provides
coverage of more websites, but at the expense of complicating the end
user experience. We might explore such techniques in Wildcard in the
future, but we think that it's valuable for end users to have a
customization experience decoupled from the challenge of web scraping
the underlying data. Sifter implements scraping across multiple pages, a
valuable feature for sorting and filtering that isn't present in
Wildcard.

Thresher \citep{hogue2005} helps end users create wrappers that map
website content to Semantic Web schemas, making customizations available
using the schema information. Wildcard shares the general idea of
wrappers with Thresher, but maps to a single generic data type, rather
than more specific schemas, increasing the range of websites and data
that Wildcard is able to support.

There are many software customization tools that offer simplified forms
of programming for end users. Chickenfoot \citep{bolin2005} and
Coscripter \citep{leshed2008} offer user friendly syntax for writing web
automation scripts; Applescript \citep{cook2007} has a similar goal for
desktop customization. There are visual programming environments for
customization that don't involve writing any text:
\href{https://support.apple.com/guide/automator/welcome/mac}{Automator}
for Mac and
\href{https://apps.apple.com/us/app/shortcuts/id915249334}{Shortcuts}
for iOS are modern options for customizing Apple products, and
\href{https://zapier.com/}{Zapier} enables users to connect different
web applications together visually. As mentioned previously, these tools
all require writing imperative programs in some form, in contrast to the
more declarative and direct approach of table-driven customization.

@@

Another relevant area involves direct manipulation interfaces for
interacting with structured data. We take inspiration from these tools
in our work, but apply them in a different domain: customizing existing
software applications, rather than interacting with databases or
constructing software from scratch.

The most closely related work is in systems that offer spreadsheet-like
querying of relational data, as proposed by Liu and Jagadish
\citep{liu2009}. SIEUFERD by Bakke and Karger \citep{bakke2016} is an
example; their paper presents a survey of other similar tools. Our work
is particularly influenced by the authors' observation that direct
manipulation requires that the user manipulate the results of a database
query rather than the query itself, and that the user must see
intermediate results at every step of constructing the query. SIEUFERD's
interface supports a far wider range of queries than Wildcard, but the
basic ideas of the user interface are similar.

@@

Our work is also inspired by the many projects that have explored using
spreadsheets as a foundation for building software applications,
including Object Spreadsheets \citep{mccutchen2016}, Quilt
\citep{benson2014}, Gneiss \citep{chang2014}, Marmite \citep{wong2007},
\href{https://www.glideapps.com/}{Glide}. These projects share the idea
that a spreadsheet is a convenient interface for editing the data
underlying a GUI application. We share that idea, but apply it to
software customization, rather than building software from scratch.

Another system in this space is ScrAPIr, by Alrashed et
al.~\citep{alrashed2020}, which enables end users to access backend web
APIs without programming. ScrAPIr shares our high level goal of end user
empowerment, as well as the idea of wrappers, by creating a shared
library of wrappers around existing web APIs. Unlike Wildcard, however,
ScrAPIr targets explicit APIs exposed by developers. It also focused on
backend services and doesn't aim to extend the frontend interfaces of
web applications.
